---
title: "Data_Tidying"
author: "Kylie Wheelock Riley"
date: "9/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(tibble.print_min = 5)
```

# Tidying Data - joins, pivots, etc. 

## Pivot_Longer
```{r pivot_longer}
##BDI spans many columns, now we are combinging all of the BDI scores into 1 column
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  pivot_longer(
    pulse_data, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    values_to = "bdi")

## The BDI values in the visit column all have a long prefix, the code below can fix this is pivot_longer
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    pulse_data,
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
    ) 

## Doing all of the cleaning in one step, also uses mutate to change the variable type to what you need 
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)
  
##Change baseline (bl) var to 00m to match other vars (subset of code above)
mutate(
  visit = recode(visit, "bl" = "00m")
)

```

## Using separate in the FAS Group dataset
```{r litters data}
## reading in data
read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>%
  count(group) 

## Tidying 
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(
    dose = str_to_lower(dose),
    wt_gain = gd18_weight - gd0_weight) %>%
  arrange(litter_number)
```

## Learning Assessment 1
 In the litters data, the variables gd0_weight and gd18_weight give the weight of the mother mouse on gestational days 0 and 18. 
Write a data cleaning chain that retains only litter_number and these columns; 
produces new variables gd and weight; 
makes gd a numeric variable taking values 0 and 18 (for the last part, you might want to use recode …). Is this version “tidy”?
```{r Learning assessment 1}
litters_data %>% 
  select(litter_number, gd0_weight, gd18_weight) %>% 
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = "gd", 
    values_to = "weight") %>% 
  mutate(gd = recode(gd, "gd0_weight" = 0, "gd18_weight" = 18))

```

# Pivot Wider
```{r pivot wider}
##This is the correct format for additional analysis or visualization, but doesn’t facilitate quick comparisons for human readers.
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)

# #An alternative presentation of the same data might have groups in rows, times in columns, and mean values in table cells.
pivot_wider(
  analysis_result, 
  names_from = "time", 
  values_from = "mean")
```

# Binding Rows
```{r}
## reading in the 3 tables in excel
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")

## tidying data and binding rows
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>%
  janitor::clean_names() %>%
  pivot_longer(
    female:male,
    names_to = "sex", 
    values_to = "words") %>%
  mutate(race = str_to_lower(race)) %>% 
  select(movie, everything()) 
```

# Joining Datsets
```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))

fas_data = 
  left_join(pup_data, litter_data, by = "litter_number")
fas_data 
```

# Learning Assessment 2
```{r}
la_2 = 
  read_csv("./data/surv_os.csv") %>%
  janitor::clean_names() %>%
  rename(uni = what_is_your_uni, os = what_operating_system_do_you_use) %>% 
  separate(uni, into = c("header", "student"), sep = "_") %>% 
  mutate(
    student = as.numeric(student)
  ) %>% 
  select(student, os, -header)



la_2_git =
  read_csv("./data/surv_program_git.csv") %>%
  janitor::clean_names() %>% 
  rename(uni = what_is_your_uni, degree = what_is_your_degree_program, git_exp = which_most_accurately_describes_your_experience_with_git) %>% 
  separate(uni, into = c("header", "student"), sep = "_") %>% 
  mutate(
    student = as.numeric(student)
  ) %>% 
  select(student, degree, git_exp, -header)

student_survey_join =
  left_join(la_2, la_2_git, by = "student")

```

